---
title: "C2: Applying Trex across cohorts"
author: 
- name: Nick Borcherding
  email: ncborch@gmail.com
  affiliation: Washington University in St. Louis, School of Medicine, St. Louis, MO, USA
date: "January 10, 2022"
output:
  BiocStyle::html_document:
    toc_float: true
---

In reference to the manuscript, this analysis centers of Figure 2 and 3


# Loading Data and Trex

Here we are using: 
* Autoencoder model in Trex, specifically for the Kidera factors. 
* The relationship between sequences being called by threshold. 
* 50 dimensions will be returned using the spectral transformation.


```{r}
list <- readRDS("./data/Fullintegrated_Tcell_seuratObjects.rds")
library(Trex)
list <- runTrex(list, 
                    edit.method = "lv",
                    AA.method = "auto", 
                    AA.properties = "KF", 
                    nearest.method = "threshold",
                    threshold = 0.85,
                    near.neighbor = 40,
                    add.INKT = FALSE,
                    add.MAIT = FALSE,
                    reduction.name = "Trex.KF")
saveRDS(list, file = "./data/Fullintegrated_Tcell_seuratObjects_wTrex.rds")
```


## Examining Trex Dimensions 

One thing I found while building this package was the interesting differences between dimensions - the first several dimensions are clearly influenced by clone, however, as the dimensions increase a more continuous-like relationship emerges.

```{r}
dir.create("./output/Figure2")
list <- readRDS("./data/Fullintegrated_Tcell_seuratObjects_wTrex.rds")
library(patchwork)
dims <- list(c(1,2), c(3,4), c(5,6), c(7,8), c(9,10))

for (i in seq_along(dims)) {
  plot <- DimPlot(list, dims = unlist(dims[i]), reduction = "Trex.KF", group.by = "CTaa") +
    scale_color_manual(values = viridis::viridis_pal(option = "B")(length(unique(list$CTaa)))) + 
    guides(color = "none") + 
    theme(plot.title = element_blank(), 
          axis.text = element_blank(), 
          axis.ticks = element_blank(),
          panel.border = element_rect(colour = "black", fill=NA, size=1), 
          axis.title = element_blank())
  assign(paste0("Plot", i), plot)
}

Plot1 + Plot2 + Plot3 + Plot4 + Plot5+  plot_layout(ncol = 5)
ggsave("./output/Figure2/AllSamples_DimExamination_clones.png", height = 3, width = 15, dpi = 600, bg = "transparent")


for (i in seq_along(dims)) {
  plot <- DimPlot(list, dims = unlist(dims[i]), reduction = "Trex.KF", group.by = "Patient") +
    scale_color_manual(values = viridis::viridis_pal(option = "H")(length(unique(list$Patient)))) + 
    guides(color = "none") + 
    theme(plot.title = element_blank(), 
          axis.text = element_blank(), 
          axis.ticks = element_blank(),
          panel.border = element_rect(colour = "black", fill=NA, size=1), 
          axis.title = element_blank())
  assign(paste0("Plot", i), plot)
}
  
  Plot1 + Plot2 + Plot3 + Plot4 + Plot5+  plot_layout(ncol = 5)
ggsave("./output/Figure2/AllSamples_DimExamination_patient.png", height = 3, width = 15, dpi = 600, bg = "transparent")

```

# Calculating UMAPS TCR vectors

```{r}
list <- RunUMAP(list, 
                        umap.method = "umap-learn",
                        #graph = "Trex.KF_snn", 
                     reduction = "Trex.KF",
                     dims = 1:30,
                        reduction.name = 'Trex.umap', 
                        reduction.key = 'trexUMAP_')
```

#Calculating and Visualizng Clonal Clusters
```{r}

############################
#Calculating clonal clusters
#############################
library(bluster)
list <- clonalCommunity(list, 
                        reduction.name = "Trex.KF", 
                        cluster.parameter = KNNGraphParam(k=100))

df <- data.frame(list@reductions$Trex.umap@cell.embeddings, list[[]])


ggplot(df, aes(x=trexUMAP_1, trexUMAP_2)) + 
  stat_density_2d(aes(fill = stat(level)),
                              geom = "polygon", 
                          bins = 20)  + 
  facet_grid(Tissue ~ ARDS) + 
  scale_fill_viridis(option = "B") + 
  theme_classic() + 
  xlim(-15,35) + 
  ylim(-15,25)
ggsave("./output/Figure2/trexDistribution.pdf", height = 6, width = 7)

df$trex.clusters <- as.character(df$trex.clusters)
count.cluster <- df %>%
  group_by(Patient, Tissue, ARDS, trex.clusters) %>%
  dplyr::count()

count.cluster <- count.cluster %>%
  group_by(Patient, Tissue) %>%
  mutate(total = sum(n)) %>%
  ungroup() %>%
  mutate(proportion = n/total)

library(stringr)
levels <- str_sort(unique(count.cluster$trex.clusters), numeric = TRUE)
count.cluster$trex.clusters <- factor(count.cluster$trex.clusters, levels = levels)

ggplot(count.cluster, aes(y = proportion, x = Tissue, fill = trex.clusters)) + 
  geom_bar(position = "fill", stat = "summary", fun.y = "mean", lwd = 0.1, color = "white") + 
  facet_grid(.~ARDS) + 
  geom_bar_text(aes(label = trex.clusters), color = "white", 
                position = "fill", stat = "summary", reflow = TRUE, 
                place = 'middle', 
                min.size = 3) + 
  theme_classic() +  
  scale_fill_viridis(option = "B", discrete = TRUE) + 
  guides(fill = "none")
ggsave("./output/Figure2/Specific.Trex.Clusters.pdf", height = 4, width = 6)
```

# Calculating UMAP for RNA and ADT

```{r}
list <- RunUMAP(list, reduction = 'pca', 
                dims = 1:30, 
                assay = 'integrated', 
              reduction.name = 'rna.umap', 
              reduction.key = 'rnaUMAP_', 
              umap.method = "umap-learn")

list <- RunUMAP(list, reduction = 'apca', 
                dims = 1:20, 
                assay = 'ADT.integrated', 
              reduction.name = 'adt.umap', 
              reduction.key = 'adtUMAP_', 
              umap.method = "umap-learn")
```



# Calculating the multi-modal neighbor 

```{r}
list <- FindMultiModalNeighbors(
  list, 
  k.nn = 40,
  reduction.list = list("pca", "apca", "Trex.KF"), 
  dims.list = list(1:30, 1:20, 1:30), modality.weight.name = "RNA.weight"
)
list <- RunUMAP(list, nn.name = "weighted.nn", 
                     reduction.name = "wnn.umap", reduction.key = "wnnUMAP_")
list <- FindClusters(list, graph.name = "wsnn", resolution = 0.6,
                          algorithm = 1, verbose = FALSE)

#WNN UMAP
dir.create("./output/Figure3")
levels <- str_sort(unique(list$trex.clusters), numeric = TRUE)
list$trex.clusters <- factor(list$trex.clusters, levels = levels)

DimPlot(list, reduction = "wnn.umap", group.by = "trex.clusters") + 
  scale_color_manual(values = viridis::viridis_pal(option = "H")(length(unique(list$trex.clusters)))) + 
  theme_void() + 
  theme(plot.title = element_blank()) + 
  
        NoLegend()
ggsave(filename = paste0("./output/Figure3/WNN.clone.umap.png"), height = 5, width = 5, dpi = 600)

plot10 <- DimPlot(list, reduction = "rna.umap", group.by = "trex.clusters") +
   guides(color = "none") + 
  scale_color_manual(values = viridis::viridis_pal(option = "H")(length(unique(list$trex.clusters)))) +
  theme_void() + 
  theme(plot.title = element_blank())
plot11 <- DimPlot(list, reduction = "adt.umap", group.by = "trex.clusters") +
   guides(color = "none") + 
  scale_color_manual(values = viridis::viridis_pal(option = "H")(length(unique(list$trex.clusters)))) +
  theme_void() + 
  theme(plot.title = element_blank())
plot12 <- DimPlot(list, reduction = "Trex.umap", group.by = "trex.clusters") +
   guides(color = "none") + 
  scale_color_manual(values = viridis::viridis_pal(option = "H")(length(unique(list$trex.clusters)))) +
  theme_void() + 
  theme(plot.title = element_blank())

plot10 + plot11 + plot12 + plot_layout(ncol=1)
ggsave(filename = paste0("./output/Figure3/ALL.umap.png"), height = 15, width = 5, dpi = 600)
```

Plotting Protein Densities

```{r}
library(Nebulosa)
DefaultAssay(list) <- "ADT"
 
genes <- c("CD45RO-TotalSeqC", "CD45RA-TotalSeqC", "CD8-TotalSeqC", "CD4-TotalSeqC", "CD25-TotalSeqC", "CTLA4-TotalSeqC", "LAG3-TotalSeqC", "PDL1-TotalSeqC")

for (i in seq_along(genes)) {
  plot_density(list, 
               reduction = "wnn.umap", 
               features =genes[i]) + 
    scale_color_viridis(option = "B") + 
    guides(color = "none") + 
    theme_void() + 
    theme(plot.title = element_blank())
  ggsave(filename = paste0("./output/Figure3/", genes[i], ".png"), height = 3, width = 3, dpi = 600)
}

saveRDS(list, file = "./data/Fullintegrated_Tcell_seuratObjects_wTrex.rds")
```

# Reducing the to clonotype representations
#Recalculates the UMAP and multimodal neighbors

```{r}
conga <- CoNGAfy(list, method = "dist", assay = c("RNA", "ADT"), 
                 meta.carry = c("CTaa", "CTgene"))

saveRDS(conga, file = "Integrated.Conga.rds")

conga <- readRDS("./data/Integrated.Conga.rds")

DefaultAssay(conga) <- "RNA"
conga <- NormalizeData(conga)
conga <- FindVariableFeatures(conga)
conga <- ScaleData(conga)
conga <- RunPCA(conga)

conga <- RunUMAP(conga, reduction = 'pca', 
                dims = 1:30, 
                assay = 'RNA', 
              reduction.name = 'rna.umap', 
              reduction.key = 'rnaUMAP_', 
              umap.method = "umap-learn")

DefaultAssay(conga) <- "ADT"
VariableFeatures(conga, assay = "ADT") <- rownames(conga[["ADT"]])
conga <- NormalizeData(conga, 
                       normalization.method = 'CLR', margin = 2)
conga <- ScaleData(conga)
conga <- RunPCA(conga, 
                reduction.name = "apca")

conga <- RunUMAP(conga, reduction = 'pca', 
                dims = 1:30, 
                assay = 'ADT', 
              reduction.name = 'ADT.umap', 
              reduction.key = 'adtUMAP_', 
              umap.method = "umap-learn")

conga <- runTrex(conga, 
                    edit.method = "lv",
                    AA.method = "auto", 
                    AA.properties = "KF", 
                    nearest.method = "threshold",
                    threshold = 0.85,
                    near.neighbor = 40,
                    add.INKT = FALSE,
                    add.MAIT = FALSE,
                    reduction.name = "Trex.KF")


 conga<- RunUMAP(conga, 
                        umap.method = "umap-learn",
                        #graph = "Trex.KF_snn", 
                     reduction = "Trex.KF",
                     dims = 1:30,
                        reduction.name = 'Trex.umap', 
                        reduction.key = 'trexUMAP_')
 saveRDS(conga, file = ".data/Integrated.Conga.rds")


 meta <- df %>% 
   group_by(CTaa, Patient, trex.clusters) %>%
   count()
x <- table(meta$CTaa)
x <- x[x > 1]

'%!in%' <- Negate("%in%")
  meta <- meta[meta$CTaa %!in% names(x), ]
  rownames.meta <- meta$CTaa
  meta <- meta[,-1]
  rownames(meta) <- rownames.meta
conga <- AddMetaData(conga, meta)  

conga$ARDS <- ifelse(conga$Patient == "S8", "moderate", "severe")
levels <- str_sort(unique(conga$trex.clusters), numeric = TRUE)
conga$trex.clusters <- factor(conga$trex.clusters, levels = levels)

#######################################################
#Remove Cluster 12, skews distribution as it is so large
######################################################

conga <- subset(conga, trex.clusters != "trex.12")
conga <- FindMultiModalNeighbors(
  conga, 
  k.nn = 40,
  reduction.list = list("pca", "apca", "Trex.KF"), 
  dims.list = list(1:30, 1:20, 1:30), modality.weight.name = "RNA.weight"
)

conga <- RunUMAP(conga, nn.name = "weighted.nn", 
                     reduction.name = "wnn.umap", reduction.key = "wnnUMAP_")

levels <- str_sort(unique(conga$trex.clusters), numeric = TRUE)
conga$trex.clusters <- factor(conga$trex.clusters, levels = levels)

conga <- FindClusters(conga, graph.name = "wknn", algorithm = 3, resolution = 0.6, verbose = FALSE)

DimPlot(conga, reduction = "wnn.umap", label = TRUE) + 
  theme_void() + 
  theme(plot.title = element_blank()) + 
  scale_colour_tableau(palette = "Tableau 20") + 
  
        NoLegend()
ggsave(filename = paste0("./output/Figure3/WNN.conga.clone.umap.png"), height = 5, width = 5, dpi = 600)
```

# Conga Protein Densities

```{r}
genes <- c("CD45RO-TotalSeqC", "CD45RA-TotalSeqC", "CD8-TotalSeqC", "CD4-TotalSeqC", "CD25-TotalSeqC", "CTLA4-TotalSeqC", "LAG3-TotalSeqC", "PDL1-TotalSeqC")

for (i in seq_along(genes)) {
  plot_density(conga, 
               reduction = "wnn.umap", 
               features =genes[i]) + 
    scale_color_viridis(option = "B") + 
    guides(color = "none") + 
    theme_void() + 
    theme(plot.title = element_blank())
  ggsave(filename = paste0("./output/Figure3/", genes[i], ".wnn.png"), height = 3, width = 3, dpi = 600)
}
```

